---
title: "Untitled"
author: "Zhenqi Wang"
date: "November 9, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
# 243 final project
```{}
install.packages(c("devtools", "roxygen2", "testthat", "knitr"))
install.packages("rstudioapi")
rstudioapi::isAvailable("0.99.149")
devtools::install_github("hadley/devtools")
library(devtools)
library(roxygen2)
has_devel()
devtools::create("~/MFA")
```

```{r}

############ BUILD A CLASS #################


# create a s4 class "mfa"
# 'mfa' has 4 attributes:
# mfa@eigenvalues: is a vector of square of singlular value(diagonal elements of delta)
# mfa@common_factor_score: is a matrix
# mfa@partial_factor_score: is a list of K matrix, K is the number of data groups
# mfa@loadings: is a matrix, the final Q

setClass(
Class="mfa",
slots=list(
weights="numeric",
eigenvalues="numeric",
common_factor_score="matrix",
partial_factor_score="list",
loadings="matrix")
)



############ constructor function: building the model #################

# constructor function: to construct 'mfa' and run the model to get attributes
# parameter: data: could be a matrix or a data.frame, should be in the same order of sets
# parameter: sets: list of vector contains vector of indices of each group
# eg. sets=c(1:3,4:5), means the 1:3 columns of data is Group 1 and the next 4:5 columns is Group2
# center and scale: the same parameters as in the function scale(), logical values or a numeric vector 
mfa<-function(data,sets,ncomps=NULL,center=TRUE,scale=TRUE){
  
  # scale and center
  data<-scale(data,center,scale)
  
  # divide data into several group according to values in sets
  # store the ith group of data to variable "Groupi"
  for (i in 1:length(sets)) {
    assign(paste0("Group",i),data.matrix(data[,min(sets[[i]]):max(sets[[i]])]))
  }
  
  
  # for each data groups conduct svd
  # store the first singular values in singularvalues
  singularvalues<-c(rep(1,dim(data)[1]))
  for (i in 1:length(sets)) {
    singularvalues[i]<-max(svd(eval(parse(text=paste0("Group",i))))$d)
  }
  
  
  # construct A to compute Q: QAQt=I
  # A's diagonal elements are the inverse of the first square singular values
  # each first square singular value is expanded to the same dimension of each data group
  expanded<-c()
  for (i in 1:length(sets)){
    expanded<-c(expanded,rep(singularvalues[i],max(sets[[i]])-min(sets[[i]])+1))
  }
  A<-diag(x = 1/expanded^2,length(expanded),length(expanded))
  #print(dim(A))
  A_half<-diag(x = 1/expanded,length(expanded),length(expanded))
  A_half_inv<-diag( x = expanded,length(expanded),length(expanded))
  
  
  
  # construct M to compute P: PMPt=I
  M<-diag(x=1/(dim(data)[1]),dim(data)[1],dim(data)[1])
  M_half<-diag(x=1/sqrt((dim(data)[1])),dim(data)[1],dim(data)[1])
  M_half_inv<-diag(x=sqrt((dim(data)[1])),dim(data)[1],dim(data)[1])
  
  
  #  X: the whole data
  X<-data.matrix(data)
  # Construct S=XAXt
  S<-as(X %*% A %*% t(X),"matrix")
  
  # do spectral decomposition on S: S=P*LAMBDA*Pt, PtMP=I
  # construct inverse delta: delta^2=LAMBDA
  eigens<-eigen(S)
  #d<-matrix(0,dim(X)[1],dim(X)[1])
  #for (i in 1:length(eigens$values)){
  #  d[i,i]<-eigens$values[i]
  #}
  d <- diag(eigens$values)
  u<-eigens$vectors
  lambda<-as(M_half %*% d %*% M_half,"matrix")
  
  delta_value<-diag(as(sqrt(lambda),"matrix"))
  
  delta_inv<-1/sqrt(lambda)
  delta_inv[is.infinite(delta_inv)]<-0
  
  # P is PMPt=I FOR S=P*LAMBDA*Pt
  P <- as(M_half_inv %*% u,"matrix")
  #print("P")
  #print(P[,1:2])
  # Q FOR Q=Xt*M*P*DELTA_inverse
  Q <- as(t(X) %*% M %*% P %*% delta_inv, "matrix")
  
  
  
  # build a list: 'partial_factor_score' to store partial factor score
  # build a matrix: 'common_factor_score' to store common factor score
  # the partial score of group i is named "Partial Score: Group i"
  # partial factor score i = no. of group * A_i* data group i* Q_i
  # common factor score = sum of partial factor score i
  
  
  partial_factor_score<-list()
  common_factor_score<-0
  for (i in 1:length(sets)){
    datai<- data.matrix(eval(parse(text=paste0("Group",i))))
    score<-length(sets) * (1/singularvalues[i]^2) * datai %*% t(datai)  %*% M %*% P %*% delta_inv
    partial_factor_score[[paste0("Partial Score: Group ",i)]]=as(score,"matrix")
    common_factor_score<-score+common_factor_score
  }
  common_factor_score<-common_factor_score/length(sets)
  
  
  # loading uses Q
  new (Class = "mfa",
    weights = diag(A),
    eigenvalues = delta_value^2,
    common_factor_score = as(common_factor_score,"matrix"),
    partial_factor_score = partial_factor_score,
    loadings = as(Q,"matrix")
  )
}

######################### test ###################
#load wine data
data<-read.csv("wine.csv",header=T,stringsAsFactors=F)
datas<-data[,2:54]
ndatas<-apply(datas,2,function(x){ (x-mean(x))/norm(x-mean(x),type="2")})
test<-mfa(ndatas,sets=list(1:6,7:12,13:18,19:23,24:29,30:34,35:38,39:44,45:49,50:53),center=FALSE,scale=FALSE)
```



```{r}
################ supplementary method #############


# set print() to print basic infomation
setMethod("print",
  signature="mfa",
  function(x,...){
    cat(paste("There are",length(x@eigenvalues),"components."),"\n")
    cat("The eigenvalue of the first component is: ",  x@eigenvalues[1],"\n")
    cat("The eigenvalue of the second component is: ",  x@eigenvalues[2],"\n")
  }
)
print(test)

# set plot() to plot table given two dimensions
#setGeneric("plot",function(object)standardGeneric("plot"))
setMethod("plot",signature="mfa",
function(object){
}
)
# set eigenvalues() to take 'mfa' and return a table (like Table 2)
setGeneric("eigenvalues",function(x,...) standardGeneric("eigenvalues"))
setMethod("eigenvalues",signature="mfa",
  function(x,...){
    eigenvalue <- x@eigenvalues
    singular_value <- sqrt(eigenvalue)
    cumulative <- cumsum(eigenvalue)
    inertia <- eigenvalue/sum(eigenvalue)*100
    cumulative_precentage <- cumulative/sum(eigenvalue)*100
    
    df <- data.frame(rbind(singular_value,eigenvalue,cumulative,inertia,cumulative_precentage))
    colnames(df) <- 1:length(eigenvalue)
    rownames(df) <- c("Singular value", "Eigenvalue","Cumulative","% Inertia","Cumulative % Inertia")
    df
  }
)
eigenvalues(test)

# set contributions() to take 'mfa' and return a matrix of contributions
setGeneric("contributions",function(x)standardGeneric("contributions"))
setMethod("contributions",signature="mfa",
  function(x){
    
    
    K <- length(x@partial_factor_score)
    q <- x@loadings
    t <- as.data.frame(table(x@weights))
    f <- x@common_factor_score
    alpha <- unique(x@weights)
    sum <- 0
    ctr_table <- NULL
    ctr_obs <- NULL
    ctr_var <- NULL
    for(i in 1:K){
      l <- t[which(t==alpha[i]),2]
      ctr_var <- rbind(ctr_var,alpha[i]*q[(sum+1):(sum+l),]^2)
      ctr_table <- rbind(ctr_table,apply(alpha[i]*q[(sum+1):(sum+l),]^2,2,sum))
      sum <- sum+l
    }
    
    lambda <- apply(1/nrow(f)*f^2,2,sum)
    ctr_obs <- sweep(1/nrow(f)*f^2,2,STATS = lambda, FUN = "/")
    list(observations=ctr_obs,
         variables=ctr_var,
         table=ctr_table)
  }
)

# set funtion RV() to take two tables and return rv coefficient
RV<-function(table1,table2){
	t1 <- t(table1)
	t2 <- t(table2)
  sum(diag((table1 %*% t1) %*% (table2 %*% t2)))/sqrt(sum(diag((table1 %*% t1) %*% (table1 %*% t1)))*sum(diag((table2 %*% t2) %*% (table2 %*% t2))))
}
RV_table <- function(dataset,sets){
  if(!is.data.frame(dataset)&!is.matrix(dataset)){stop("dataset must be a matrix or a dataframe")}
    RV <- matrix(NA,length(sets),length(sets))
    for(i in 1:length(sets)){
      for(j in i:length(sets)){
        rv <- RV(as.matrix(dataset[,sets[[i]]]),as.matrix(dataset[,sets[[j]]]))
        RV[i,j] <- rv
        RV[j,i] <- rv
      }
    }
    RV
}

RV_table(ndatas,sets=list(1:6,7:12,13:18,19:23,24:29,30:34,35:38,39:44,45:49,50:53))


# set funtion LG() to take two tables and return lg coefficient
LG<-function(table1,table2){
	t1 <- t(table1)
	t2 <- t(table2)
	gamma1<-svd(table1)$d[1]
	gamma2<-svd(table2)$d[1]
  sum(diag((table1 %*% t1) %*% (table2 %*% t2)))/(gamma1^2*gamma2^2)
}


LG_table <- function(dataset,sets){
  if(!is.data.frame(dataset)&!is.matrix(dataset)){stop("dataset must be a matrix or a dataframe")}
    LG <- matrix(NA,length(sets),length(sets))
    for(i in 1:length(sets)){
      for(j in i:length(sets)){
        lg <- LG(as.matrix(dataset[,sets[[i]]]),as.matrix(dataset[,sets[[j]]]))
        LG[i,j] <- lg
        LG[j,i] <- lg
      }
    }
    LG
}
LG_table(ndatas,sets=list(1:6,7:12,13:18,19:23,24:29,30:34,35:38,39:44,45:49,50:53))

# Bootstrap?
```



########################### Shiny ###########################

# for shiny
library(shiny)
ui <- fluidPage()
server <- function(input, output) {
  # for inputs
  name<-reactive({
    startpoint<-c(input$x,input$y)
  })
    
  output$plot <- renderPlot({
    
    # create a S4 class 
  
    # initialize use constructor function
    
    # set new method 
    
    # instanced
    data=read.csv("")
    my_mfa<-mfa()
    plot(mfa)
  })
}

ui<-fluidPage(
  titlePanel(h1("MFA")),
  plotOutput("plot"),
    fluidRow(
)


shinyApp(ui = ui, server = server)


